### 2019-06-27 report/chapter04.tex
Methods update
Results update

% LaTeX file for Chapter 04
<<'preamble04',include=FALSE>>=
library(knitr)
opts_chunk$set(
    fig.path='figure/ch04_fig', 
    self.contained=FALSE,
    cache=FALSE
) 
@

\chapter{Methods}
\section{Introduction and Notation}
The analysis has already been said to be restricted on clinical or health care interventions. The interventions are restricted to two participant arms of which some measure of sanity or worsening of health is measured and compared. The difference between these two is referred to as the effect of the treatment. Where it is not particularly mentioned, the term treatment effect refers to any effect measure such as log risk ratio, log hazard ratio, Cohen's $d$ or Hedges $g$, fisher's z score or pearson correlation. \\
Let us consider a meta-analysis with $n$ study treatment effects ($n > 1$, but typically small). A study is indexed by $i$, and it's treatment effect by  $\theta_i$. The observed treatment effect is $\hat{\theta}_i$. The pooled treatment effect of a meta-analysis will be denoted as $\theta_M$, and consequently, the observed pooled treatment effect as $\hat{\theta}_M$. Furthermore, each treatment effect is typically measured with some standard error $s_i$ and an estimate of $s_i$ is given by $\hat{s}_i$. \\
For continuous outcomes, let $m_t$ be the mean of the treatment group, $m_c$ the mean of the control group, and equivalentyl $sd_t$ and $sd_c$ the corresponding standard deviations. $n_t$ and $n_c$ are the total number of participants in the groups.
In the case of binary outcomes, let $e_t$ be the count of events in the treatment arm $e_c$ the count of events in the control group. The observed counts in a study $i$ are referred to as $e_{t,i}$ and analogously $e_{c,i}$.

\section{Effect Measures and p-values}
\subsection{Continuous Outcomes}
For given $(m_t, m_c), (sd_t, sd_c)$ and $(n_t, n_c)$, one can compute mean difference as well as a standardized mean difference (here: Hedges $g$) and  a standard error thereof. Note that the definition of Hedges $g$ and its standard error $s$ varies amongst the literature, the following applies for this report:

\begin{align}
s &= \sqrt{\frac{(n_t - 1)sd_t^2 + (n_c - 1)sd_c^2}{n_t + n_c - 2}} & g &= \frac{m_t - m_c}{s} \label{eq:hedges.g}
\end{align}

@@

\section{Small study effect and Publication Bias Adjustment}
There are different approaches to correct for small study effects and publication bias. They can mainly be distinguished by their underlying methods: regression based approaches aim to regress the effect to a study with infinte precision (i.e. very small standard error) or to a summary effect, corrected for publication bias. Selection modells aim to simulate different, hypothetical selection processes and attain a approximate treatment effect by sensitivity analysis:

@@

$\chi^2$ can be used with a $\chi^2$ distribution with one degree of freedom to obtain a $p$-value. Note that the test is almost equivalent to Egger's small study effect test with $\tau^2 = 0$. Thus, although the copas selection model models publication bias, it is dependent on the small study effects to find the most suitable pair $(a,b)$. \\
If one applies the model to a single meta-analysis, a sensitivity analysis is suggested. One can observe how $\theta$ and it's confidence intervals change dependent on the underlying selection process. Selection models are in general not recommended for inference \citep{selection.assessment}. 
\citet{limitmeta} have shown how the method can be implemented in a simulation for inference purposes. \\
A range of values of $(a,b)$ are applied, and the test for residual small study effect as described in equation \ref{eq:copas.small.study} is applied. If all obtained $p$-values from the test are above a threshold 0.1, this is interpreted as no evidence, and no need for modelling, and the standard, classical random effects meta-analysis is retained. If none of the $p$-values is above the threshold, a wider range of values for $(a,b)$ is used. When some $p$-values are above, and some below the threshold, the pair $(a,b)$ with the smallest number of missing studies is retained (that is, the least intense underlying selection model is chosen).\\
Currently, there is no test to detect missspecifications in the model itself, the authors themselves have argued that a nonparametric test of the residuals would lack power.

\section{Transformation between effect sizes} \label{sec:transformation.effectsizes}
Assumon that binary outcomes result from a dichtomozation of originally continuous random variables, in this case, the logistic distribution, a transformation from typical binary effect measures to Cohen's $d$ can be achieved (\citet{Intro.meta}, pg. 47). \\
Let $\theta$ be a log odds ratio and $s$ its standard error. Cohen's $d$ and it's variance $s_d^2$ is obtained by

@@

$\frac{\pi}{\sqrt{3}} = 1.81$ is the standard deviation of the logistic distribution $L(\mu, \eta)$ with scale parameter $\eta = 1$, so we just divide the log odds ratio and it's variance throught the standard deviation. The approximation works only well if $e_t$ and $e_c$ are not very small, especially in the case of $s_d^2$. Plugging in the observed log odds ratio $\hat{\theta}$ and $\hat{s}^2$ will give an estimate of Cohen's $d$.\\
Pearsons correlation can be attained by the formulas (\citet{olkin1985dtor}, \citet{Intro.meta} pg. 48-49)

@@

% If $\beta_{0} \ne 0$, the null hypothesis of no small study effect may be contested, using that $\beta_{0} \sim t_{n-1}$, $n-1$ being the degrees of freedom of the $t$-distribution. The p-value for $\beta_{0} = 0$ (no reporting bias) is then given by \ref{egger.pvalue}.

@@

% If we accept $\beta = 0$, then we accept that the selection model has satisfactorily explained any relationship between $y$ and $s$. Only if the value is large enough, typically $p > 0.05$, one concludes that the selection model has explained the observed data. The p-value is obtained by a likelihood ratio test comparing the maximum of the likelihood with the $\beta$ term added and without it, and by a likelihood ratio test.

@@

% In practice, only a range of values for $a,b$ are reasonable. For those values, the quantities above can be calculated and illustrated. Values for $\mu$ that have p-values over a predefined significance threshold can be used for inference of the effect size. 

>>>>

% LaTeX file for Chapter 04

\chapter{Methods} 
\section{Introduction and Notation}
The analysis has already been said to be restricted on clinical or health care interventions. The interventions are restricted to comparisons of two treatment arms by some measure of sanity or worsening of health. The difference in this measure between the grouops is referred to as the treatment effect. Where it is not particularly mentioned, the term treatment effect refers to any effect measure such as log risk ratio, log hazard ratio, Cohen's $d$ or Hedges $g$, Fisher's z score or Pearsons correlation coefficient. \\
Let us consider a meta-analysis with $n$ study treatment effects ($n > 1$, but typically small). A study is indexed by $i$, and it's treatment effect by  $\theta_i$. The observed treatment effect is $\hat{\theta}_i$. The pooled treatment effect of a meta-analysis will be denoted as $\theta_M$, and consequently, the observed pooled treatment effect as $\hat{\theta}_M$. Furthermore, each treatment effect is typically measured with some standard error $s_i$ and an estimate of $s_i$ is denoted as $\hat{s}_i$. The $\hat{}$ sign thus indicates if it is an estimate.\\
For continuous outcomes, let $m_t$ be the mean of the treatment group, $m_c$ the mean of the control group, and equivalently $\textrm{sd}_t$ and $\textrm{sd}_c$ the corresponding standard deviations. $n_t$ and $n_c$ are the total number of participants in the groups.
In the case of binary outcomes, let $e_t$ be the count of events in the treatment arm $e_c$ the count of events in the control group. The observed counts in a study $i$ are referred to as $e_{t,i}$ and analogously $e_{c,i}$.

\section{Effect Measures and $p$-values}
\subsection{Continuous Outcomes}
For given $(m_t, m_c), (\textrm{sd}_t, \textrm{sd}_c)$ and $(n_t, n_c)$, one can compute mean difference as well as a standardized mean difference (here: Hedges $g$) and  a standard error thereof. Note that the definition of Hedges $g$ and its standard error $s$ varies among the literature, the following applies for this report:

\begin{align}
s &= \sqrt{\frac{(n_t - 1)\textrm{sd}_t^2 + (n_c - 1)\textrm{sd}_c^2}{n_t + n_c - 2}} & g &= \frac{m_t - m_c}{s} \label{eq:hedges.g}
\end{align}

@@

\section{Small Study Effect and Publication Bias Adjustment}
There are different approaches to correct for small study effects and publication bias. They can mainly be distinguished by their underlying methods: regression based approaches aim to regress the effect to a study with infinite precision (i.e. very small standard error) or to a summary effect, corrected for publication bias. Selection models aim to simulate different, hypothetical selection processes and attain a approximate treatment effect by sensitivity analysis:

@@

$\chi^2$ can be used with a $\chi^2$ distribution with one degree of freedom to obtain a $p$-value. Note that the test is almost equivalent to Egger's small study effect test with $\tau^2 = 0$. Thus, although the copas selection model models publication bias, it is dependent on the small study effects to find the most suitable pair $(a,b)$. \\
If one applies the model to a single meta-analysis, a sensitivity analysis is suggested. One can observe how $\theta$ and it's confidence intervals change dependent on the underlying selection process. Selection models are in general not recommended for inference (e.g. \citet{selection.assessment}). 
\citet{limitmeta} have shown how the method can be implemented in a simulation for inference purposes. \\
A range of values of $(a,b)$ are applied, and the test for residual small study effect as described in equation \ref{eq:copas.small.study} is applied. If all obtained $p$-values from the test are above a threshold 0.1, this is interpreted as no evidence, and no need for modelling, and the standard, classical random effects meta-analysis is retained. If none of the $p$-values is above the threshold, a wider range of values for $(a,b)$ is used. When some $p$-values are above, and some below the threshold, the pair $(a,b)$ with the smallest number of missing studies is retained (that is, the least intense underlying selection model is chosen).\\
Currently, there is no test to detect miss-specifications in the model itself, the authors themselves have argued that a non-parametric test of the residuals would lack power.

\section{Transformation between effect sizes} \label{sec:transformation.effectsizes}
Assuming that binary outcomes result from a dichotomization of originally continuous random variables, in this case, the logistic distribution, a transformation from typical binary effect measures to Cohen's $d$ can be achieved \cite[47]{Intro.meta}. \\
Let $\theta$ be a log odds ratio and $s$ its standard error. Cohen's $d$ and it's variance $s_d^2$ is obtained by

@@

$\frac{\pi}{\sqrt{3}} = 1.81$ is the standard deviation of the logistic distribution $L(\mu, \eta)$ with scale parameter $\eta = 1$, so we just divide the log odds ratio and it's variance through the standard deviation. The approximation works only well if $e_t$ and $e_c$ are not very small, especially in the case of $s_d^2$. Plugging in the observed log odds ratio $\hat{\theta}$ and $\hat{s}^2$ will give an estimate of Cohen's $d$.\\
Pearson's correlation can be attained by the formulas (\citet{olkin1985dtor}, \cite[48]{Intro.meta})

@@

% If $\beta_{0} \ne 0$, the null hypothesis of no small study effect may be contested, using that $\beta_{0} \sim t_{n-1}$, $n-1$ being the degrees of freedom of the $t$-distribution. The $p$-value for $\beta_{0} = 0$ (no reporting bias) is then given by \ref{egger.pvalue}.

@@

% If we accept $\beta = 0$, then we accept that the selection model has satisfactorily explained any relationship between $y$ and $s$. Only if the value is large enough, typically $p > 0.05$, one concludes that the selection model has explained the observed data. The $p$-value is obtained by a likelihood ratio test comparing the maximum of the likelihood with the $\beta$ term added and without it, and by a likelihood ratio test.

@@

% In practice, only a range of values for $a,b$ are reasonable. For those values, the quantities above can be calculated and illustrated. Values for $\mu$ that have $p$-values over a predefined significance threshold can be used for inference of the effect size. 
